---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Nora Jones 
    url: https://example.com/norajones
    affiliation: Spacely Sprockets
    affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


This is the OLD RV Sparse Code:


## RV Sparse

An *RV Sparse* estimator consists of calculating the realized variance but with a sample that is less frequent than the 1-minute grid. Instead, it is sampled in an *s*-minute grid (where $s \geq 1$).

Here, we will use different combinations of $s$ such as `s = 5`, `s = 10`, `s = 15`, and `s = 30`. In order to sample in an *s*-minute grid, we make use of the following auxiliary rolling-functions:

```{r}
# For a 5-minute grid:
five_roll_sum <- slidify(
  .f = ~sum(.x), 
  .period = 5, 
  .align = "right")

# For a 10-minute grid:
ten_roll_sum <- slidify(
  .f = ~sum(.x), 
  .period = 10, 
  .align = "right")


# For a 15-minute grid:
fifteen_roll_sum <- slidify(
  .f = ~sum(.x), 
  .period = 15, 
  .align = "right")


# For a 30-minute grid:
thirty_roll_sum <- slidify(
  .f = ~sum(.x), 
  .period = 30, 
  .align = "right")
```

We can now subset the data to a five-minute grid. We'll call this new data `IPC_5m`.

```{r}
IPC_5m <- IPC_fr %>% 
  mutate(ret_5m = five_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_5m) %>%
  filter(!is.na(ret_5m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 5))
```

```{r}
RV_s5 <- IPC_5m %>% 
  mutate(log_ret_sq = ret_5m^2) %>% 
  group_by(index_date) %>% 
  summarise(RVs5 = sum(log_ret_sq),
            RVs5_an = 252 * sqrt(RVs5))
```

Now we repeat the process for a $s = 10$ grid:

```{r}
IPC_10m <- IPC_fr %>%
  mutate(ret_10m = ten_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_10m) %>%
  filter(!is.na(ret_10m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 10))
```

```{r}
RV_s10 <- IPC_10m %>% 
  mutate(log_ret_sq = ret_10m^2) %>% 
  group_by(index_date) %>% 
  summarise(RVs10 = sum(log_ret_sq),
            RVs10_an = 252 * sqrt(RVs10))
```

And then again, for the $s = 15$ grid:

```{r}
IPC_15m <- IPC_fr %>% 
  mutate(ret_15m = fifteen_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_15m) %>%
  filter(!is.na(ret_15m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 15))
```

```{r}
RV_s15 <- IPC_15m %>% 
  mutate(log_ret_sq = ret_15m^2) %>% 
  group_by(index_date) %>% 
  summarise(RVs15 = sum(log_ret_sq),
            RVs15_an = 252 * sqrt(RVs15))
```

And lastly, for the $s = 30$ grid:

```{r}
IPC_30m <- IPC_fr %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_30m) %>%
  filter(!is.na(ret_30m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 30))
```

```{r}
RV_s30 <- IPC_30m %>% 
  mutate(log_ret_sq = ret_30m^2) %>% 
  group_by(index_date) %>% 
  summarise(RVs30 = sum(log_ret_sq),
            RVs30_an = 252 * sqrt(RVs30))
```

