---
title: "Realized Higher Moment Measures"
description: |
  Exploring Realized Higher Moment Measures on the S&P/BMV IPC Index
author:
  - name: Alexis Solis Cancino
    affiliation: ITAM
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_depth: 2
    css: "D:/11-Estancia/Finance-Thesis/05-Resources/styles.css"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  
  # Places figures on their own pages
  fig.pos = 'p',
  
  # Figure resolution and size
  out.width = '100%',
  dpi = 400,
  
  # Latex figure environment
  fig.env = "figure"
)


# --- Load the libraries! ---

# Data Wrangling
library(tidyverse)
library(lubridate)
library(skimr)
# library(tidylog)

# Time Series
library(timetk)
library(tidyquant)
library(tibbletime)
library(tsibble)
library(hms)
library(modeltime)
library(tidymodels)

# Visualization
library(ggthemes)
library(scales)
library(ghibli)
library(cowplot)
library(plotly)
library(gt)

# Reproducibility
library(usethis)
library(here)


# Speed
library(data.table)
library(dtplyr)
library(arrow)


# --- Setup the ggplot theme ---
source(here("05-Resources", "ggplot2_themes.R"))
```

```{r}
# Read the data
IPC_fr <- read_parquet(file = here("01-Data", "parquet", "IPC_fr.parquet"))
```

Problems with the data:

1.  We don't have the same number of datapoints (trades) per day.

```{r}
IPC_fr %>%
  count(tidy_date) %>% 
  arrange(n)
  # ggplot(aes(x = tidy_date, y = n)) +
  # geom_col() +
  # theme_roboto()
```

2.  We don't have the same number of datapoints (trades) per time (i.e. hour:minute).

```{r}
IPC_fr %>% count(tidy_time) %>% 
  arrange(n)
#   ggplot(aes(x = tidy_time, y = n)) +
#   geom_col() +
#   theme_roboto()
```

# Realized Variance

For this section, we'll transform our dataset into a `timetibble` object, so that time filtering and wrangling of the data is easier. We'll call the new dataset `IPC_tbl_tm` (IPC tibble time).

```{r}
IPC_tbl_tm <- IPC_fr %>%  
  tibbletime::as_tbl_time(index = tidy_date)

IPC_tbl_tm
```

We can now compute the realized variance for each day:

```{r}
RVs <- IPC_tbl_tm %>% 
  mutate(log_ret_sq = log_ret^2) %>% 
  group_by(tidy_date) %>% 
  summarise(RV = sum(log_ret_sq),
            RV_an = 252 * sqrt(RV)) %>% 
  
  # Get rid of date "2001-09-12", which has a 0 RV.
  filter(!RV == 0)

RVs
```

Now let's plot the RVs:

```{r, layout="l-body-outset", fig.width=8, fig.asp=0.618}
rv_plot <- RVs %>% 
  # filter_time("1996" ~ "end") %>%
  ggplot(mapping = aes(x = tidy_date, y = RV_an)) +
  geom_point(color = amazing_colors[1], alpha = 0.5) +
  # geom_smooth(method = "loess") +
  labs(
    title = "Intraday Volatility for the S&P/BMV IPC Index",
    subtitle = "As measured by the annualized daily-realized-variance",
    x = NULL,
    y = "Annualized RV"
    ) +
  scale_y_continuous(labels = scales::number_format()) +
  scale_x_date(date_labels = "%Y", date_breaks = "2 years") +
  geom_smooth(colour = amazing_colors[2])

rv_plot
```

```{r}
# ggsave(
#   filename = "signature-plot.png",
#   units = "in",
#   width = 8 * 1.61803398875,
#   height = 8,
#   path = "D:/01-Data-Science/01-RStudio-Projects/Finance-Research/03_images/",
#   dpi = 400,
#   device = "png"
# )
```

## RV Sparse

An *RV Sparse* estimator consists of calculating the realized variance but with a sample that is less frequent than the 1-minute grid. Instead, it is sampled in an *s*-minute grid (where $s \geq 1$).

Here, we will use different combinations of $s$ such as `s = 5`, `s = 10`, `s = 15`, and `s = 30`. In order to sample in an *s*-minute grid, we make use of the following auxiliary rolling-functions:

```{r}
# # For a 5-minute grid:
# five_rolling_sum <- rollify(.f = ~sum(.x), window = 5)
# 
# # For a 10-minute grid:
# ten_rolling_sum <- rollify(.f = ~sum(.x), window = 10)
# 
# # For a 15-minute grid:
# fifteen_rolling_sum <- rollify(.f = ~sum(.x), window = 15)
# 
# # For a 30-minute grid:
# thirty_rolling_sum <- rollify(.f = ~sum(.x), window = 30)
```

```{r}
# For a 5-minute grid:
five_roll_sum <- slidify(
  .f = ~sum(.x), 
  .period = 5, 
  .align = "right")

# For a 10-minute grid:
ten_roll_sum <- slidify(
  .f = ~sum(.x), 
  .period = 10, 
  .align = "right")


# For a 15-minute grid:
fifteen_roll_sum <- slidify(
  .f = ~sum(.x), 
  .period = 15, 
  .align = "right")



# For a 30-minute grid:
thirty_roll_sum <- slidify(
  .f = ~sum(.x), 
  .period = 30, 
  .align = "right")
```

We can now subset the data to a five-minute grid. We'll call this new data `IPC_5m`.

```{r}
IPC_5m <- IPC_tbl_tm %>% 
  mutate(ret_5m = five_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_5m) %>%
  filter(!is.na(ret_5m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 5))
```

```{r}
RV_s5 <- IPC_5m %>% 
  mutate(log_ret_sq = ret_5m^2) %>% 
  group_by(tidy_date) %>% 
  summarise(RVs5 = sum(log_ret_sq),
            RVs5_an = 252 * sqrt(RVs5))
```

Now we repeat the process for a $s = 10$ grid:

```{r}
IPC_10m <- IPC_tbl_tm %>%
  mutate(ret_10m = ten_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_10m) %>%
  filter(!is.na(ret_10m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 10))
```

```{r}
RV_s10 <- IPC_10m %>% 
  mutate(log_ret_sq = ret_10m^2) %>% 
  group_by(tidy_date) %>% 
  summarise(RVs10 = sum(log_ret_sq),
            RVs10_an = 252 * sqrt(RVs10))
```

And then again, for the $s = 15$ grid:

```{r}
IPC_15m <- IPC_tbl_tm %>% 
  mutate(ret_15m = fifteen_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_15m) %>%
  filter(!is.na(ret_15m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 15))
```

```{r}
RV_s15 <- IPC_15m %>% 
  mutate(log_ret_sq = ret_15m^2) %>% 
  group_by(tidy_date) %>% 
  summarise(RVs15 = sum(log_ret_sq),
            RVs15_an = 252 * sqrt(RVs15))
```

And lastly, for the $s = 30$ grid:

```{r}
IPC_30m <- IPC_tbl_tm %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_30m) %>%
  filter(!is.na(ret_30m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 30))
```

```{r}
RV_s30 <- IPC_30m %>% 
  mutate(log_ret_sq = ret_30m^2) %>% 
  group_by(tidy_date) %>% 
  summarise(RVs30 = sum(log_ret_sq),
            RVs30_an = 252 * sqrt(RVs30))
```

## Volatility signature plot

The Sparse RV Estimator is calculated using an *s*-minute grid ($s \geq 1$):

\\begin{equation} RV\\\_{t+1}\^s = \\sum \\limits\_{j=1}\^{m/s} R\^2\_{t+js/m} \\end{equation}

Now the problem lies in the choice of the subsampling parameter $s$, a problem which represents a typical bias-variance trade-off. That is because ...

This is where volatility signature plots (VSP) will help; these are a graphical tool that helps us choose the subsampling parameter $s$. We do a VSP in the following manner:

1.  Compute $RV_{t+1}^s$ for values of $s$ going from 1 to 120 minutes.

```{r}
all_RV_sparse <- RVs %>% 
  left_join(y = RV_s5, by = "tidy_date") %>% 
  left_join(y = RV_s10, by = "tidy_date") %>% 
  left_join(y = RV_s15, by = "tidy_date") %>% 
  left_join(y = RV_s30, by = "tidy_date")

all_RV_sparse
```

2.  Scatter plot the average RV across days on the vertical axis against $s$ on the horizontal axis.

```{r}
signature_plot_data <- all_RV_sparse %>% 
  summarise(s1 = mean(RV),
            s5 = mean(RVs5),
            s10 = mean(RVs10),
            s15 = mean(RVs15),
            s30 = mean(RVs30)
            ) %>% 
  pivot_longer(cols = everything(), names_to = "s", values_to = "avg_RV_sparse") %>% 
  add_column(minutes = c(1, 5, 10, 15, 30))
```

```{r, layout="l-body-outset", fig.width=8, fig.asp=0.618}
signature_plot_data %>% 
  ggplot(aes(minutes, avg_RV_sparse)) +
  # geom_smooth(colour = amazing_colors[2]) +
  geom_point(colour = amazing_colors[3], size = 3) +
  labs(
    title = "Signature Plot for the RV Sparse",
    x = "Minutes",
    y = "RV"
  ) +
  scale_y_continuous(labels = number_format())
```

3.  Look for the smallest $s$ such that the average RV does not change much for values of $s$ larger than this number.

## Average RV Estimator

We can also sample with an *s*-minute grid, but every *k* minutes. Here, we employ several combinations of these: make use of `s = 5` and `s = 10` minute samples every `k = 1` minute, and `s = 15` + `s = 30` minute samples every `k = 5` minutes. We also try `s = 15` every `k = 1` minute and `s = 10` every `k = 2`.

5-minute grid every 1-minute:

```{r}
IPC_5m_1m <- IPC_tbl_tm %>% 
  mutate(ret_5m = five_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_5m) %>%
  filter(!is.na(ret_5m))


IPC_5m_1m
```

10-minute grid every 1-minute:

```{r}
IPC_10m_1m <- IPC_tbl_tm %>% 
  mutate(ret_10m = ten_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_10m) %>%
  filter(!is.na(ret_10m))

IPC_10m_1m
```

10-minute grid every 2-minutes:

```{r}
IPC_10m_2m <- IPC_tbl_tm %>% 
  mutate(ret_10m = ten_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_10m) %>%
  filter(!is.na(ret_10m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 2))

IPC_10m_2m
```

15-minute grid every 1-minute:

```{r}
IPC_15m_1m <- IPC_tbl_tm %>% 
  mutate(ret_15m = fifteen_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_15m) %>%
  filter(!is.na(ret_15m))

IPC_15m_1m
```

15-minute grid every 3-minutes:

```{r}
IPC_15m_3m <- IPC_tbl_tm %>% 
  mutate(ret_15m = fifteen_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_15m) %>%
  filter(!is.na(ret_15m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 3))

IPC_15m_3m
```

15-minute grid every 5-minutes:

```{r}
IPC_15m_5m <- IPC_tbl_tm %>% 
  mutate(ret_15m = fifteen_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_15m) %>%
  filter(!is.na(ret_15m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 5))

IPC_15m_5m
```

30-minute grid every 1-minute:

```{r}
IPC_30m_1m <- IPC_tbl_tm %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_30m) %>%
  filter(!is.na(ret_30m))

IPC_30m_1m
```

30-minute grid every 3-minutes:

```{r}
IPC_30m_3m <- IPC_tbl_tm %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_30m) %>%
  filter(!is.na(ret_30m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 3))

IPC_30m_3m
```

30-minute grid every 5-minutes:

```{r}
IPC_30m_5m <- IPC_tbl_tm %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_30m) %>%
  filter(!is.na(ret_30m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 5))

IPC_30m_5m
```

30-minute grid every 10-minutes:

```{r}
IPC_30m_10m <- IPC_tbl_tm %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(tidy_date, tidy_time, ret_30m) %>%
  filter(!is.na(ret_30m)) %>% 
  slice(seq(1, nrow(IPC_tbl_tm), by = 10))

IPC_30m_10m
```

Now, let's compute the RVs for each s-minute every k-minute grid:

```{r}
IPC_5m_1m %>% 
  mutate(log_ret_sq = ret_5m^2) %>% 
  group_by(tidy_date) %>% 
  summarise(RV = sum(log_ret_sq),
            RV_an = 252 * sqrt(RV))
```
