---
title: "Realized Higher Moment Measures"
description: |
  Exploring Realized Higher Moment Measures on the S&P/BMV IPC Index
author:
  - name: Alexis Solis Cancino
    affiliation: ITAM
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_depth: 2
    css: "D:/11-Estancia/Finance-Thesis/05-Resources/styles.css"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  
  # Places figures on their own pages
  fig.pos = 'p',
  
  # Figure resolution and size
  out.width = '100%',
  dpi = 400,
  
  # Latex figure environment
  fig.env = "figure"
)


# --- Load the libraries! ---

# Data Wrangling
library(tidyverse)
library(lubridate)
library(skimr)
library(glue)
# library(tidylog)

# Time Series
library(timetk)
library(tidyquant)
# library(hms)
# library(modeltime)
# library(tidymodels)

# Visualization
# library(ggthemes)
library(scales)
# library(ghibli)
library(cowplot)
# library(plotly)
library(gt)
library(progress)

# Reproducibility
library(usethis)
library(here)


# Speed
# library(data.table)
# library(dtplyr)
library(arrow)
library(furrr)


# Setup parallelization plan
future::plan(strategy = multiprocess, workers = 11)

# --- Setup the ggplot theme ---
source(here("05-Resources", "ggplot2_themes.R"))
```

```{r}
# Read the data
IPC_fr <- read_parquet(file = here("01-Data", "parquet", "IPC_fr2.parquet"))
```

# Realized Volatility

To estimate daily volatility of equity returns, we can use the intraday data, in particular, the intraday returns. Since the expected value of the intraday returns is zero, a good estimator of variance is 

```{r}
# IPC_fr <- IPC_fr %>%  
  # tibbletime::as_tbl_time(index = index_date)

# IPC_fr
```

We can now compute the realized variance for each day:

```{r}
RVs <- IPC_fr %>% 
  mutate(log_ret_sq = log_ret^2) %>% 
  group_by(index_date) %>% 
  summarise(RV = sum(log_ret_sq),
            RV_an = sqrt(252 * RV),
            .groups = "drop") %>% 
  
  # Get rid of date "2001-09-12", which has a 0 RV.
  filter(!RV == 0)

RVs
```

Now let's plot the RVs:

```{r, layout="l-body-outset", fig.width=8, fig.asp=0.618}
rv_plot <- RVs %>% 
  # filter_time("1996" ~ "end") %>%
  ggplot(mapping = aes(x = index_date, y = RV)) +
  geom_col(color = amazing_colors[8]) +
  labs(
    title = "Intraday Variance for the S&P/BMV IPC Index",
    subtitle = "Daily Realized Variance",
    x = NULL,
    y = NULL
    ) +
  scale_y_continuous(labels = scales::number_format()) +
                     # limits = c(0, 0.005)) +
  scale_x_date(date_labels = "%Y", date_breaks = "2 years")
  # geom_smooth(colour = amazing_colors[2])

rv_plot
```




```{r}
RVs %>% 
    filter(RV > 0.001)
```




```{r}
# ggsave(
#   filename = "signature-plot.png",
#   units = "in",
#   width = 8 * 1.61803398875,
#   height = 8,
#   path = "D:/01-Data-Science/01-RStudio-Projects/Finance-Research/03_images/",
#   dpi = 400,
#   device = "png"
# )
```


```{r}
RVs %>% 
  pull(RV) %>% 
  acf(lag.max = 100, plot = F) %>% 
  broom::tidy() %>% 
  filter(lag != 0) %>% 
  ggplot(aes(lag, acf)) +
  geom_line()
```



## RV Sparse

An *RV Sparse* estimator consists of calculating the realized variance but with a sample that is less frequent than the 1-minute grid. Instead, it is sampled in an *s*-minute grid (where $s \geq 1$).

Here, we will use different combinations of $s$ such as `s = 5`, `s = 10`, `s = 15`, and `s = 30`. In order to plot a *signature plot*, we will also sample from $s \in [1, 120]$.



```{r}
# Define progress bar
pb <- progress::progress_bar$new(
  format = " computing RVs [:bar] :percent eta: :eta",
  total = 100,
  clear = FALSE
)
```



```{r}
get_RV_sparse <- function(.sparse, 
                          .ret_tbl = IPC_fr, 
                          .index_by = index_date) {

  # Print progress bar
  # pb$tick()
  
  # Define rolling sum function
  roll_sum <- slidify(
    .f = ~sum(.x), 
    .period = .sparse, 
    .align = "right")
  
  # Define new return table
  new_ret_tbl <- .ret_tbl %>% 
    mutate(ret_grid = roll_sum(log_ret)) %>% 
    select({{ .index_by }}, ret_grid) %>%
    filter(!is.na(ret_grid)) %>% 
    slice(seq(1, nrow(.ret_tbl), by = .sparse))
  
  # Define new RV table
  new_RV_tbl <- new_ret_tbl %>% 
    mutate(log_ret_sq = ret_grid^2) %>% 
    group_by({{ .index_by }}) %>% 
    summarise(
      RV_sparse = sum(log_ret_sq),
      RV_sparse_an = sqrt(252 * RV_sparse),
      .groups = "drop"
      # "RVs{{ .sparse }}" := sum(log_ret_sq),
      # "RVs{{ .sparse }}_an" := 252 * sqrt(eval(as.name(paste0("RVs", {{.sparse}}))))
      )
  
  return(new_RV_tbl)
  
}
```

## Volatility signature plot

The Sparse RV Estimator is calculated using an *s*-minute grid ($s \geq 1$):

$$ 
RV_{t+1}^s = \sum \limits_{j=1}^{m/s} R^2_{t+js/m} 
$$

```{r}
get_RV_sparse(.sparse = 5)
```


Now the problem lies in the choice of the sub-sampling parameter $s$, a problem which represents a typical bias-variance trade-off. That is because ...

This is where *volatility signature plots* (VSP) will help; these are a graphical tool that helps us choose the sub-sampling parameter $s$. We do a VSP in the following manner:

1.  Compute $RV_{t+1}^s$ for values of $s$ going from 1 to 120 minutes.

```{r get-all-RV-sparse, cache=T}
# Define a vector with all the "s-grid" parameters
sparse_grid <- seq(from = 1, 
                   to = 240, 
                   by = 1)


# Create a tibble with all the nested data
nested_RV_sparse <- tibble(
  my_result = sparse_grid %>% future_map(.f = ~ get_RV_sparse(.sparse = .)),
  s_grid = sparse_grid
)
```

```{r unnest-RV-sparse}
# Expand the tibble
all_RV_sparse <- nested_RV_sparse %>% unnest(cols = c(my_result))


# Get the mean for each RV sparse:
RV_means <- all_RV_sparse %>% 
  group_by(s_grid) %>% 
  summarise(
    mean_RV_sparse = mean(RV_sparse),
    .groups = "drop"
  )
```

2.  Scatter plot the average RV across days on the vertical axis against $s$ on the horizontal axis.

```{r, layout="l-body-outset", fig.width=8, fig.asp=0.618}
RV_means %>% 
  ggplot(aes(s_grid , mean_RV_sparse)) +
  geom_point(alpha = 0.6, 
             colour = amazing_colors[1], 
             size = 2) +
  labs(
    title = "Signature Plot for the Annualized RV Sparse",
    x = "Minutes",
    y = "RV"
  ) +
  scale_y_continuous(labels = number_format())
```

3.  Look for the smallest $s$ such that the average RV does not change much for values of $s$ larger than this number.

## Average RV Estimator

We can also sample with an *s*-minute grid, but every *k* minutes. Here, we employ several combinations of these: make use of `s = 5` and `s = 10` minute samples every `k = 1` minute, and `s = 15` + `s = 30` minute samples every `k = 5` minutes. We also try `s = 15` every `k = 1` minute and `s = 10` every `k = 2`.

5-minute grid every 1-minute:

```{r}
IPC_5m_1m <- IPC_fr %>% 
  mutate(ret_5m = five_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_5m) %>%
  filter(!is.na(ret_5m))


IPC_5m_1m
```

10-minute grid every 1-minute:

```{r}
IPC_10m_1m <- IPC_fr %>% 
  mutate(ret_10m = ten_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_10m) %>%
  filter(!is.na(ret_10m))

IPC_10m_1m
```

10-minute grid every 2-minutes:

```{r}
IPC_10m_2m <- IPC_fr %>% 
  mutate(ret_10m = ten_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_10m) %>%
  filter(!is.na(ret_10m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 2))

IPC_10m_2m
```

15-minute grid every 1-minute:

```{r}
IPC_15m_1m <- IPC_fr %>% 
  mutate(ret_15m = fifteen_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_15m) %>%
  filter(!is.na(ret_15m))

IPC_15m_1m
```

15-minute grid every 3-minutes:

```{r}
IPC_15m_3m <- IPC_fr %>% 
  mutate(ret_15m = fifteen_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_15m) %>%
  filter(!is.na(ret_15m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 3))

IPC_15m_3m
```

15-minute grid every 5-minutes:

```{r}
IPC_15m_5m <- IPC_fr %>% 
  mutate(ret_15m = fifteen_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_15m) %>%
  filter(!is.na(ret_15m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 5))

IPC_15m_5m
```

30-minute grid every 1-minute:

```{r}
IPC_30m_1m <- IPC_fr %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_30m) %>%
  filter(!is.na(ret_30m))

IPC_30m_1m
```

30-minute grid every 3-minutes:

```{r}
IPC_30m_3m <- IPC_fr %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_30m) %>%
  filter(!is.na(ret_30m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 3))

IPC_30m_3m
```

30-minute grid every 5-minutes:

```{r}
IPC_30m_5m <- IPC_fr %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_30m) %>%
  filter(!is.na(ret_30m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 5))

IPC_30m_5m
```

30-minute grid every 10-minutes:

```{r}
IPC_30m_10m <- IPC_fr %>% 
  mutate(ret_30m = thirty_roll_sum(log_ret)) %>% 
  select(index_date, index_time, ret_30m) %>%
  filter(!is.na(ret_30m)) %>% 
  slice(seq(1, nrow(IPC_fr), by = 10))

IPC_30m_10m
```

Now, let's compute the RVs for each s-minute every k-minute grid:

```{r}
IPC_5m_1m %>% 
  mutate(log_ret_sq = ret_5m^2) %>% 
  group_by(index_date) %>% 
  summarise(RV = sum(log_ret_sq),
            RV_an = 252 * sqrt(RV))
```
